---
title: "Macroeconometrics: An investigation of the effects of financial conditions on the real economy"
author: 
  - name: "Thomas Kronholm Moeller"
# format:
#   pdf:
#     fontfamily: cmbright
#     geometry: margin=1.5cm
format:
  html:
    toc: true
    toc-location: left
---

> Disclaimer: This page is an ongoing research project conducted as a part of Macroeconometrics (ECOM9007) at The University of Melbourne, Australia.\

## Research question & Motivation
As the financial crises emerged in 2008, the world economy realized just how much financial conditions affects the real economy. This research project intends to quantify the effects of a tightening of financial conditions on the real economy. 

The objective question to be answered can thus be summarized as follows; Does a tightening of financial conditions have the anticipated effects on the real economy, and if so, what are the magnitudes reasonable.       

**What motivates this?**

It is or rather has been a standard part of macroeconomic modelling to exclude the financial sector from applied theoretical modelling. This has been a popular choice of researchers, who has argued that there is no or at least negligible relationship between financial and real variables. Nevertheless, financial crisis are often followed by significant drops in consumption and production, which might be a result of lower consumer sentiment, affecting the overall demand for goods as well as savings. It is therefore of importance for various policy makers to apply these insight when forming the economic policy as tighter monetary and/or macroprudential stances might have some undesired outcomes. 

Furthermore, the research question is highly applicable in today's economic climate as central banks has started raising rates world wide, the financial condition index enables us to get a complete view of the affects that this might have by looking at financial markets, credit and liquidity while including the shadow banking system.          

## Data and its properties 
As financial variables in general are fast moving compared to real economic variables, one obstacle was to find economic variables, which was able to mimic the effects on the real economy although on a higher frequency. 

The economic variables used for the empirical analysis are as follows: 

The economic activity is modeled by the **industrial production** which is an approximate variable mimicing the movements in Gross Domestic Product (GDP). Furthermore, as the United States are considered a closed economy, industrial production is considered sufficient.

Additionally, the overall **consumer price index** has been introduced to the model as well, as financial conditions might affect the inflation through various expectations. 

In order to directly link the movements in the financial condition index and the real economy total **lending** from commercial banks is applied. 

As aforementioned, one could hypothesize that the effect which financial conditions has on the real economy was through consumer expectations. Therefore, the model includes **consumer sentiment** provided by University of Michigan. 

Lastly, as we are interested in identifying a shock to financial conditions, the **National Financial Condition Index (NFCI)** is included. The index is constructed by the Federal Reserve, Chicago and is a measure of the conditions of finance, taking financial markets, credit and liquidity and the shadow banking system into account. 

The time series are retrieved using the database of Fred, Economic data provided by the Federal Reserve Bank og St.Louis, and is available on a monthly basis. The time period considered is from 01.01.1990 - 01.12.2022. The data is obatained using package **fredr()** 

```{r Getting data from fred}
#| echo: false
#| message: false
#| warning: false

library(fredr)
library(ggplot2)
library(dplyr)
library(purrr)
library(vars)
library(xts)
library(gridExtra)
library(tseries)
library(tidyverse)
library(palmerpenguins)
library(quarto)
library(tinytex)
#Setting up API key from FRED
fredr_set_key("54706a95d44824ac499f1012d9b3a401")

start_date <- as.Date("1990-01-01")
end_date <- as.Date("2022-12-01")

# Define a vector of series IDs
#timeseries <- c("INDPRO", "CPIAUCSL", "UMCSENT", "TOTCI", "USSTHPI", "NFCI") incl. house prices
timeseries <- c("INDPRO", "CPIAUCSL", "UMCSENT", "TOTCI", "NFCI")
data <- list()

# Retrieve the data for the series and date range ::::: change m to q for quarterly
for (name in timeseries) {
  fred_data <- fredr(series_id = name, observation_start = start_date, observation_end = end_date, frequency = "m")
  data[[name]] <- fred_data
}

#Deleting columns not required 
indp <- data[["INDPRO"]]
indp$ln_value <- log(indp$value)
indp_ <- indp[, c("date","ln_value")]
indp_$date <- as.Date(indp_$date, format = "%d-%m-%Y")
tindp <- xts(indp_$ln_value, order.by = indp_$date)

cpi <- data[["CPIAUCSL"]]
cpi$ln_value <- log(cpi$value)
cpi_ <- cpi[, c("date","ln_value")] 
tcpi <- xts(cpi_$ln_value, order.by = indp_$date)

expec <- data[["UMCSENT"]]
expec_ <- expec[, c("date","value")]
texpec <- xts(expec_$value, order.by = indp_$date)

lend <- data[["TOTCI"]]
lend$ln_value <- log(lend$value)
lend_ <- lend[, c("date","ln_value")]
tlend <- xts(lend_$ln_value, order.by = indp_$date)

#hp <- data[["USSTHPI"]]
#hp$ln_value <- log(hp$value)
#hp_ <- hp[, c("date","ln_value")]
#thp <- xts(hp_$ln_value, order.by = indp_$date)

fci <- data[["NFCI"]]
fci_ <- fci[, c("date","value")]
tfci <- xts(fci_$value, order.by = indp_$date)

#Merging the series into vector Y
#Y = na.omit(merge(tindp, tcpi, texpec, tlend, thp, tfci))
#colnames(Y)<- c("indu", "cpi", "exp", "lend","hp","fci")
Y = na.omit(merge(tindp, tcpi, texpec, tlend, tfci))
colnames(Y)<- c("indu", "cpi", "exp", "lend","fci")

```

```{r plotting the series}
#| echo: false
#| message: false
#| warning: false
#| results: hide
#Plotting the series in levels
ip = ggplot(data = Y[,1], aes(x = index(Y[,1]), y = Y[,1])) +
  geom_line(color = "black") +
  labs(title = "Industrial production", x = "Year", y = "Ln(Indp)") +
  theme_minimal()

inf = ggplot(data = Y[,2], aes(x = index(Y[,2]), y = Y[,2])) +
  geom_line(color = "black") +
  labs(title = "Consumer price index", x = "Year", y = "Ln(CPI)") +
  theme_minimal()

ex = ggplot(data = Y[,3], aes(x = index(Y[,3]), y = Y[,3])) +
  geom_line(color = "black") +
  labs(title = "Consumer Expectations", x = "Year", y = "Index") +
  theme_minimal()
    
le = ggplot(data = Y[,4], aes(x = index(Y[,4]), y = Y[,4])) +
  geom_line(color = "black") +
  labs(title = "Lending", x = "Year", y = "Ln(Lending)") +
  theme_minimal()

#hou = ggplot(data = Y[,5], aes(x = index(Y[,5]), y = Y[,5])) +
#  geom_line(color = "black") +
#  labs(title = "House price index", x = "Year", y = "Ln(HP)") +
#  theme_minimal()

fincon = ggplot(data = Y[,5], aes(x = index(Y[,5]), y = Y[,5])) +
  geom_line(color = "black") +
  labs(title = "National Financial Condition index", x = "Year", y = "Index") +
  theme_minimal()


```

**Proterties of the data**
The five time series are presented in the figure below. 

```{r showing plot}
#| echo: false
#| #| message: false
#| warning: false
#grid.arrange(ip, inf, ex, le, hou, fincon, nrow = 3, ncol = 2)
grid.arrange(ip, inf, ex, le, fincon, nrow = 3, ncol = 2)

```

```{r test}
#| echo: false
#Doing an ACF test :::::::: change to 6 if HP incl.
adf_ <- list()
for (i in 1:5) {
  adf_result = adf.test(Y[,i])
  adf_[[i]] <- adf_result
}
head(adf_)

# Testing for cointegration rank with trend
vecm_Y = ca.jo(Y, type = "trace", ecdet = "trend", K = 5, spec = "transitory")
summary(vecm_Y) 

# Testing for cointegration rank with constant
vecm_Y = ca.jo(Y, type = "trace", ecdet = "cons", K = 5, spec = "transitory")
summary(vecm_Y) 
```






